---
title: "exercise-05"
output: html_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
### Load in Libraries
```{r}
library(tidyverse)
library(mosaic)
library(car)
library(manipulate)
```
## Challenge 1

### Read in file
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv"
d <- read_csv(f)
```

### Filter Movies from 1920 to 1979 and that are between 1 and 3 hours long
### Add "decade" column
```{r}
d <- d %>% 
  subset(startYear >= 1920) %>%
  subset(startYear <= 1979) %>%
  subset(runtimeMinutes >= 60) %>%
  subset(runtimeMinutes <= 180) %>% 
  mutate(decade = ifelse(startYear >= 1970, "70s",
                  ifelse(startYear >= 1960, "60s",
                  ifelse(startYear >= 1950, "50s",
                  ifelse(startYear >= 1940, "40s",
                  ifelse(startYear >= 1930, "30s", "20s"))))))
```

### Make a histogram of run times for each decade
```{r}
hist <- ggplot(d, aes(runtimeMinutes)) + geom_histogram() + facet_wrap(.~decade)
hist
```

### Calculate Mean and SD of run times for each decade
```{r}
results <- data.frame(c(unique(d$decade)), c(aggregate(d$runtimeMinutes ~ d$decade, d, "mean")), c(aggregate(d$runtimeMinutes ~ d$decade, d, "sd")))   %>%
  subset(select = -c(d.decade,d.decade.1)) %>%
  setNames(c('Decade', 'Runtime_Mean (Min)', 'Runtime_SD (Min)')) 

results

# Check mean and standard variation
# r20s <- d %>% subset(decade == "20s")
# mean20s <- mean(r20s$runtimeMinutes)
# sd20s <- sd(r20s$runtimeMinutes)
```

### Create a Sample Mean of Run Times of 100 movies for Each Decade
```{r}
n <- 100 #number of samples
s <- slice_sample(d, n = n, by = decade, weight_by = NULL, replace = FALSE)

smean20s <- s %>%
  subset(decade == "20s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()

smean30s <- s %>%
  subset(decade == "30s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

smean40s <- s %>%
  subset(decade == "40s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

smean50s <- s %>%
  subset(decade == "50s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

smean60s <- s %>%
  subset(decade == "60s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

smean70s <- s %>%
  subset(decade == "70s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

Sample_Means <- c(smean20s, smean30s, smean40s, smean50s, smean60s, smean70s)

```

### Calculate Standard Deviation of Means
```{r}
# Use "s" from previous sampling
ssd20s <- s %>%
  subset(decade == "20s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()

ssd30s <- s %>%
  subset(decade == "30s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()

ssd40s <- s %>%
  subset(decade == "40s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()

ssd50s <- s %>%
  subset(decade == "50s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()

ssd60s <- s %>%
  subset(decade == "60s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()

ssd70s <- s %>%
  subset(decade == "70s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()

Sample_SDs <- c(ssd20s, ssd30s, ssd40s, ssd50s, ssd60s, ssd70s)

```

### Calculate Means from Population
```{r}
mean20s <- d %>%
  subset(decade == "20s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

mean30s <- d %>%
  subset(decade == "30s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

mean40s <- d %>%
  subset(decade == "40s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

mean50s <- d %>%
  subset(decade == "50s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

mean60s <- d %>%
  subset(decade == "60s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

mean70s <- d %>%
  subset(decade == "70s") %>%
  summarise(mean(runtimeMinutes)) %>%
  as.numeric()

Pop_Means <- c(mean20s, mean30s, mean40s, mean50s, mean60s, mean70s)
```

### Calculate Standard Error of Samples
```{r}
sSE20s <- ssd20s/sqrt(n)

sSE30s <- ssd30s/sqrt(n)

sSE40s <- ssd40s/sqrt(n)

sSE50s <- ssd50s/sqrt(n)

sSE60s <- ssd60s/sqrt(n)

sSE70s <- ssd70s/sqrt(n)

Sample_Standard_Errors <- c(sSE20s, sSE30s, sSE40s, sSE50s, sSE60s, sSE70s)
```

### Compare Sample and Population Statistics
```{r}
Decade <- c("20s", "30s", "40s", "50s", "60s", "70s")

rtStats <- data.frame(Decade, Sample_Means, Pop_Means, Sample_SDs, Sample_Standard_Errors)

rtStats
```

### Generate a Sampling Distribution of 1000 Samples
### of 100 Movies of Each Decade
```{r}
reps <- 1000

mdist20s <- vector(length = reps)
mdist30s <- vector(length = reps)
mdist40s <- vector(length = reps)
mdist50s <- vector(length = reps)
mdist60s <- vector(length = reps)
mdist70s <- vector(length = reps)

sdist20s <- vector(length = reps)
sdist30s <- vector(length = reps)
sdist40s <- vector(length = reps)
sdist50s <- vector(length = reps)
sdist60s <- vector(length = reps)
sdist70s <- vector(length = reps)
 
for (i in 1:reps) {
  samp <- slice_sample(d, n = n, by = decade, weight_by = NULL, replace = FALSE)
  mdist20s[[i]] <- samp %>%
  subset(decade == "20s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()
  
  mdist30s[[i]] <- samp %>%
  subset(decade == "30s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()
  
  mdist40s[[i]] <- samp %>%
  subset(decade == "40s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()
  
  mdist50s[[i]] <- samp %>%
  subset(decade == "50s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()
  
  mdist60s[[i]] <- samp %>%
  subset(decade == "60s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()
  
  mdist70s[[i]] <- samp %>%
  subset(decade == "70s") %>%
  summarize(mean(runtimeMinutes)) %>%
  as.numeric()
  
  sdist20s[[i]] <- samp %>%
  subset(decade == "20s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()
  
  sdist30s[[i]] <- samp %>%
  subset(decade == "30s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()
  
  sdist40s[[i]] <- samp %>%
  subset(decade == "40s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()
  
  sdist50s[[i]] <- samp %>%
  subset(decade == "50s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()
  
  sdist60s[[i]] <- samp %>%
  subset(decade == "60s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()
  
  sdist70s[[i]] <- samp %>%
  subset(decade == "70s") %>%
  summarise(sd(runtimeMinutes)) %>%
  as.numeric()
}
```

### Calculate Mean and SD of Sampling Distribution of Each Decade
```{r}
m20sDist <- mean(mdist20s)
m30sDist <- mean(mdist30s)
m40sDist <- mean(mdist40s)
m50sDist <- mean(mdist50s)
m60sDist <- mean(mdist60s)
m70sDist <- mean(mdist70s)

Sampling_Distribution_Means <- c(m20sDist, m30sDist, m40sDist, m50sDist, m60sDist, m70sDist)

sd20sDist <- mean(sdist20s)
sd30sDist <- mean(sdist30s)
sd40sDist <- mean(sdist40s)
sd50sDist <- mean(sdist50s)
sd60sDist <- mean(sdist60s)
sd70sDist <- mean(sdist70s)

Sampling_Distribution_SDs <- c(sd20sDist, sd30sDist, sd40sDist, sd50sDist, sd60sDist, sd70sDist)
```

### Plot Histogram of the Means of Each Decade
##### They Have a Normal Distribution
```{r}
hist(mdist20s)
hist(mdist30s)
hist(mdist40s)
hist(mdist50s)
hist(mdist60s)
hist(mdist70s)
```

### Compare Means of First Sample, Population, and Sampling Distribution
### for Each Decade
```{r}
rtMeanResults <- data.frame(Decade, Sample_Means, Pop_Means, Sampling_Distribution_Means)

rtMeanResults
```

## Challenge 2
### Read in File
```{r}
g <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/zombies.csv"

z <- read_csv(g)
```

### Calculate Population Mean and SD for 
### Each Quantitative Variable
```{r}
Quantitative_Variables <- c("Height", "Weight", "Age", "Zombies Killed", "Years of Education")

# Means
zHeightMean <- mean(z$height)
zWeightMean <- mean(z$weight)
zAgeMean <- mean(z$age)
zZKilledMean <- mean(z$zombies_killed)
zYearsEduMean <- mean(z$years_of_education)

Population_Means <- c(zHeightMean,zWeightMean,zAgeMean,zZKilledMean,zYearsEduMean)

# Population SD
n <- nrow(z) #Population Size

zHeightSD <- sqrt((sum((z$height - zHeightMean)^2))/n)
zWeightSD <- sqrt((sum((z$weight - zWeightMean)^2))/n)
zAgeSD <- sqrt((sum((z$age - zAgeMean)^2))/n)
zZKilledSD <- sqrt((sum((z$zombies_killed - zZKilledMean)^2))/n)
zYearsEduSD <- sqrt((sum((z$years_of_education - zYearsEduMean)^2))/n)

Population_SDs <- c(zHeightSD,zWeightSD,zAgeSD,zZKilledSD,zYearsEduSD)

PopStats <- data.frame(Quantitative_Variables,Population_Means,Population_SDs)
PopStats
```

### Make Boxplots of Each of These Variables by Gender
```{r}
zHeightBP <- ggplot(z, aes(x = gender, y = height)) +
            geom_boxplot() + xlab("Gender") + ylab("Height")
zWeightBP <- ggplot(z, aes(x = gender, y = weight)) +
            geom_boxplot() + xlab("Gender") + ylab("Weight")
zAgeBP <- ggplot(z, aes(x = gender, y = age)) +
            geom_boxplot() + xlab("Gender") + ylab("Age")
zZKilledBP <- ggplot(z, aes(x = gender, y = zombies_killed)) +
            geom_boxplot() + xlab("Gender") + ylab("Zombies Killed")
zYearsEduBP <- ggplot(z, aes(x = gender, y = years_of_education)) +
            geom_boxplot() + xlab("Gender") + ylab("Years of Education")
zHeightBP
zWeightBP
zAgeBP
zZKilledBP
zYearsEduBP
```

### Make Scatterplots of Height and Weight vs Age,
### Separated by Gender
##### For both variables males appear to have higher values,
##### though only height appears to have a positive correlation between
##### with age, whereas weight appears to have no correlation.
```{r}
sctrH <- ggplot(z, aes(x = age, y = height, color = gender)) +
  geom_point()

sctrW <- ggplot(z, aes(x = age, y = weight, color = gender)) +
  geom_point()

sctrH
sctrW
```

### Histograms and QQ Plots of Quantitative Variables
##### Height, Weight, and Age appear to be normally distributed,
##### whereas the number of zombies killed and years of 
##### education appear to be a positively skewed distribution.
```{r}
hist(z$height, xlab = "Height", ylab = "Frequency", main = "Height Histogram")
hist(z$weight, xlab = "Weight", ylab = "Frequency", main = "Weight Histogram")
hist(z$age, xlab = "Age", ylab = "Frequency", main = "Age Histogram")
hist(z$zombies_killed, xlab = "Zombies Killed", ylab = "Frequency", main = "Zombies Killed Histogram")
hist(z$years_of_education, xlab = "Years of Education", ylab = "Frequency", main = "Years of Education Histogram")

qqPlot(z$height, ylab = "Height", xlab = "Norm Quantiles", main = "Height Q-Q Plot")
qqPlot(z$weight, ylab = "Weight", xlab = "Norm Quantiles", main = "Weight Q-Q Plot")
qqPlot(z$age, ylab = "Age", xlab = "Norm Quantiles", main = "Age Q-Q Plot")
qqPlot(z$zombies_killed, ylab = "Zombies Killed", xlab = "Norm Quantiles", main = "Zombies Killed Q-Q Plot")
qqPlot(z$years_of_education, ylab = "Years of Education", xlab = "Norm Quantiles", main = "Years of Education Q-Q Plot")
```

### Sample of 50 Survivors without replacement
### Calculate Sample Means and SDs of Quantitative Variables
```{r}
sampPop <- 50
zsamp <- slice_sample(z, n = sampPop, weight_by = NULL, replace = FALSE)

sampHM <- zsamp %>%
  summarize(mean(height)) %>%
  as.numeric()
sampWM <- zsamp %>%
  summarize(mean(weight)) %>%
  as.numeric()
sampAM <- zsamp %>%
  summarize(mean(age)) %>%
  as.numeric()
sampZKM <- zsamp %>%
  summarize(mean(zombies_killed)) %>%
  as.numeric()
sampYEM <- zsamp %>%
  summarize(mean(years_of_education)) %>%
  as.numeric()

zSampMeans <- c(sampHM, sampWM, sampAM, sampZKM, sampYEM)

sampHSD <- zsamp %>%
  summarize(sd(height)) %>%
  as.numeric()
sampWSD <- zsamp %>%
  summarize(sd(weight)) %>%
  as.numeric()
sampASD <- zsamp %>%
  summarize(sd(age)) %>%
  as.numeric()
sampZKSD <- zsamp %>%
  summarize(sd(zombies_killed)) %>%
  as.numeric()
sampYESD <- zsamp %>%
  summarize(sd(years_of_education)) %>%
  as.numeric()

zSampSDs <- c(sampHSD, sampWSD, sampASD, sampZKSD, sampYESD)
```

### Calculate Standard Error Estimate and
### 95% Confidence Interval for Each Mean
```{r}
sampHSE <- sampHSD/sqrt(sampPop)
sampWSE <- sampWSD/sqrt(sampPop)
sampASE <- sampASD/sqrt(sampPop)
sampZKSE <- sampZKSD/sqrt(sampPop)
sampYESE <- sampYESD/sqrt(sampPop)

zSampSEs <- c(sampHSE, sampWSE, sampASE, sampZKSE, sampYESE) 

lowerSampHCI95 <- signif(sampHM + qnorm(p = .025) * sampHSE, digits = 5)
upperSampHCI95 <- signif(sampHM + qnorm(p = .975) * sampHSE, digits = 5)
ci95SampH <- paste(lowerSampHCI95, "-", upperSampHCI95)

lowerSampWCI95 <- signif(sampWM + qnorm(p = .025) * sampWSE, digits = 5)
upperSampWCI95 <- signif(sampWM + qnorm(p = .975) * sampWSE, digits = 5)
ci95SampW <- paste(lowerSampWCI95, "-", upperSampWCI95)

lowerSampACI95 <- signif(sampAM + qnorm(p = .025) * sampASE, digits = 5)
upperSampACI95 <- signif(sampAM + qnorm(p = .975) * sampASE, digits = 5)
ci95SampA <- paste(lowerSampACI95, "-", upperSampACI95)

lowerSampZKCI95 <- signif(sampZKM + qnorm(p = .025) * sampZKSE, digits = 5)
upperSampZKCI95 <- signif(sampZKM + qnorm(p = .975) * sampZKSE, digits = 5)
ci95SampZK <- paste(lowerSampZKCI95, "-", upperSampZKCI95)

lowerSampYECI95 <- signif(sampYEM + qnorm(p = .025) * sampYESE, digits = 5)
upperSampYECI95 <- signif(sampYEM + qnorm(p = .975) * sampYESE, digits = 5)
ci95SampYE <- paste(lowerSampYECI95, "-", upperSampYECI95)

zSampCI95s <- c(ci95SampH, ci95SampW, ci95SampA, ci95SampZK, ci95SampYE)
```

### Draw 199 Random Samples of 50 Survivors and Gather Means 
### of Quantitative Variables
```{r}
zreps <- 199

zDistH <- vector(length = zreps)
zDistW <- vector(length = zreps)
zDistA <- vector(length = zreps)
zDistZK <- vector(length = zreps)
zDistYE <- vector(length = zreps)

for (i in 1:zreps) {
  zsamp <- slice_sample(z, n = sampPop, weight_by = NULL, replace = FALSE)
  
zDistH[[i]] <- zsamp %>%
  summarize(mean(height)) %>%
  as.numeric()
  
zDistW[[i]] <- zsamp %>%
  summarize(mean(weight)) %>%
  as.numeric()

zDistA[[i]] <- zsamp %>%
  summarize(mean(age)) %>%
  as.numeric()

zDistZK[[i]] <- zsamp %>%
  summarize(mean(zombies_killed)) %>%
  as.numeric()

zDistYE[[i]] <- zsamp %>%
  summarize(mean(years_of_education)) %>%
  as.numeric()
}
```

### Add Previous Sample to the New 199 Samples
```{r}
zDistH <- append(zDistH, sampHM)
zDistW <- append(zDistW, sampWM)
zDistA <- append(zDistA, sampAM)
zDistZK <- append(zDistZK, sampZKM)
zDistYE <- append(zDistYE, sampYEM)
```

### Mean and SD of Each Variable
```{r}
zDistHM <- mean(zDistH)
zDistWM <- mean(zDistW)
zDistAM <- mean(zDistA)
zDistZKM <- mean(zDistZK)
zDistYEM <- mean(zDistYE)

zDistMs <- c(zDistHM, zDistWM, zDistAM, zDistZKM, zDistYEM)

zDistHSD <- sd(zDistH)
zDistWSD <- sd(zDistW)
zDistASD <- sd(zDistA)
zDistZKSD <- sd(zDistZK)
zDistYESD <- sd(zDistYE)

zDistSDs <- c(zDistHSD, zDistWSD, zDistASD, zDistZKSD, zDistYESD)
```

### Compare Means and SDs from Sampling Distribution to the 
### Standard Error of the First Sample
##### Distribution Standard Deviations are About Equal to the Standard
##### Errors of the First Sample
```{r}
zDistStats <- data.frame(Quantitative_Variables, zDistMs, zDistSDs, zSampSEs)
colnames(zDistStats) <- c("Variable", "Distribution Means", "Distribution Standard Deviations", "Sample Standard Errors")
zDistStats
```

### Plot Distribution Means
##### All variables now have a normal distribution, including number of
##### zombies killed and years of education, which were previously
##### positively skewed
```{r}
hist(zDistH, xlab = "Height Means", main = "Sampling Distribution Height Means")
hist(zDistW, xlab = "Weight Means", main = "Sampling Distribution Weight Means")
hist(zDistA, xlab = "Age Means", main = "Sampling Distribution Age Means")
hist(zDistZK, xlab = "Zombies Killed Means", main = "Sampling Distribution Zombies Killed Means")
hist(zDistYE, xlab = "Years of Education Means", main = "Sampling Distribution Years of Education Means")
```

### Construct 95% Confidence Interval for Each Variable 
### from Sampling Distribution
##### Confidence intervals were relatively close between the 
##### sampling distribution and the single sample
```{r}
lowerDistH <- signif(quantile(zDistH, .025), digits = 5)
upperDistH <- signif(quantile(zDistH, .975), digits = 5)
zDistCIH <- paste(lowerDistH, "-", upperDistH)

lowerDistW <- signif(quantile(zDistW, .025), digits = 5)
upperDistW <- signif(quantile(zDistW, .975), digits = 5)
zDistCIW <- paste(lowerDistW, "-", upperDistW)

lowerDistA <- signif(quantile(zDistA, .025), digits = 5)
upperDistA <- signif(quantile(zDistA, .975), digits = 5)
zDistCIA <- paste(lowerDistA, "-", upperDistA)

lowerDistZK <- signif(quantile(zDistZK, .025), digits = 5)
upperDistZK <- signif(quantile(zDistZK, .975), digits = 5)
zDistCIZK <- paste(lowerDistZK, "-", upperDistZK)

lowerDistYE <- signif(quantile(zDistYE, .025), digits = 5)
upperDistYE <- signif(quantile(zDistYE, .975), digits = 5)
zDistCIYE <- paste(lowerDistYE, "-",  upperDistYE)

zDistCIs <- c(zDistCIH, zDistCIW, zDistCIA, zDistCIZK, zDistCIYE)

DistCIs <- data.frame(Quantitative_Variables, zDistCIs, zSampCI95s)
colnames(DistCIs) <- c("Variable", "Distribution 95% CIs", "Sample 95% CIs")
DistCIs
```

### Generate a Bootstrap Distribution and Calculate 95% CI
```{r}
n_boot <- 1000

BootH <- vector(length = n_boot)
BootW <- vector(length = n_boot)
BootA <- vector(length = n_boot)
BootZK <- vector(length = n_boot)
BootYE <- vector(length = n_boot)

for (i in 1:n_boot) {
  zBoot <- slice_sample(z, n = sampPop, weight_by = NULL, replace = TRUE)
  
BootH[[i]] <- zBoot %>%
  summarize(mean(height)) %>%
  as.numeric()

BootW[[i]] <- zBoot %>%
  summarize(mean(weight)) %>%
  as.numeric()

BootA[[i]] <- zBoot %>%
  summarize(mean(age)) %>%
  as.numeric()

BootZK[[i]] <- zBoot %>%
  summarize(mean(zombies_killed)) %>%
  as.numeric()

BootYE[[i]] <- zBoot %>%
  summarize(mean(years_of_education)) %>%
  as.numeric()
}

lowerBootH <- signif(quantile(BootH, .025), digits = 5)
upperBootH <- signif(quantile(BootH, .975), digits = 5)
zBootCIH <- paste(lowerBootH, "-", upperBootH)

lowerBootW <- signif(quantile(BootW, .025), digits = 5)
upperBootW <- signif(quantile(BootW, .975), digits = 5)
zBootCIW <- paste(lowerBootW, "-", upperBootW)

lowerBootA <- signif(quantile(BootA, .025), digits = 5)
upperBootA <- signif(quantile(BootA, .975), digits = 5)
zBootCIA <- paste(lowerBootA, "-", upperBootA)

lowerBootZK <- signif(quantile(BootZK, .025), digits = 5)
upperBootZK <- signif(quantile(BootZK, .975), digits = 5)
zBootCIZK <- paste(lowerBootZK, "-", upperBootZK)

lowerBootYE <- signif(quantile(BootYE, .025), digits = 5)
upperBootYE <- signif(quantile(BootYE, .975), digits = 5)
zBootCIYE <- paste(lowerBootYE, "-",  upperBootYE)

zBootCIs <- c(zBootCIH, zBootCIW, zBootCIA, zBootCIZK, zBootCIYE)

BootCIs <- data.frame(Quantitative_Variables, zDistCIs, zSampCI95s, zBootCIs)
colnames(BootCIs) <- c("Variable", "Distribution 95% CIs", "Sample 95% CIs", "Bootstrap 95% CIs")
BootCIs
```
